Contention Management for 5G

### 2017-02-08 article/article.tex
dessins + courbes pour grande tailles
améliorations définitions, exemple de non monotonie

	We consider a directed graph $G=(V,A)$ modelling a network. Each arc  $(u,v)$ in $A$ is labeled by an integer $Dl(u,v) \geq 1$ that we call the delay and
	which represents the number of time slots taken by a signal to go from $u$ to $v$ using this arc. Note that for any arc $(u,v)$, $Dl(u,v)=Dl(v,u)$.

@@

      A {\bf routing function} $\cal R$ in $G$ associates to each pair of vertices $(u,v)$ a route from $u$ to $v$. Let $\cal C$ be an {\bf assignment} in $G$, i.e., a set of couples of different vertices of $G$. We denote by $\cal R_{\cal C}$ the set of routes ${\cal R}(u,v)$ for any $(u,v)$ in $\cal C$. $\cal R_{\cal C}$ is called an {\bf assigned graph}, and contains all the information needed in forthcoming problems (couples of vertices, routes and delay of the arcs).

      \todo{Si on s'en sert, ajouter ici que le routage est cohérent.}

   \subsection{Slotted time Model}
      Consider now a positive integer $P$ called the {\bf period}. In our problem the messages we send in the network will be periodic of period $P$ and thus we consider slices of time of $P$ discrete slots. Assume we send a message at the source of the route $r$,
      at the time slot $m$ in the first period, then a message will be sent at time $m$ at each new period. We define the first time slot at which a message reaches a vertex $v$ in this route by $t(v,r) = m + \lambda(v,r) \mod P$.

      A message usually cannot be transported in a single time slot. We denote by $\tau$ the number 
      of consecutive slots necessary to transmit a message. Let us call $[t(v,r)]$ the set of time slots used by $r$ at a vertex $v$ in a period $P$, that is $[t(v,r)] = \{t(v,r) + i \mod P \mid 0 \leq i \leq \tau \}$.

      A {\bf $P$-periodic affectation} of $\cal R_{\cal C}$ consists in a set  ${\cal M}=(m_0, \ldots ,m_{c-1})$ of $c$ integers that we call {\bf offsets}, with $c$ the cardinal of the assignment $\cal C$. The number $m_i$ represents the index of the first slot used in a period  by the route $r_i \in {\cal R}_{\cal C}$ at its source.
      A $P$-periodic affectation must have no {\bf collision} between two routes in ${\cal R}_{\cal C}$, that is $\forall (r_i, r_j) \in {\cal R}_{\cal C}^2, i \neq j$, % with $\tau$ the size (in number of consecutive slots) of each message that must be periodically sent on each route of ${\cal R}_{\cal C}$, 
      we have $$[t(u,r_i)] \cap [t(u,r_j)] = \emptyset .$$

      Notice that the notion of $P$-periodic affectation \textbf{is not monotone} with regard to $P$. 
      Indeed, consider messages of size 1, we can build a graph, with $c$ routes $r_i \in {\cal R}_{\cal C}, i < c$ which all intersect two by two and
      such that if $r_i$ and $r_j$ have $v$ as first common vertex we have $\lambda(v,r_i) - \lambda(v,r_j)=1$.
      Therefore there is a $2$-periodic affectation by setting all $m_i$ to $0$.

%       A {\bf conflict graph} represents the collision between the routes of $ {\cal R}_{\cal C}$. The vertices of a conflict graph $G = (V,E)$ are the routes of $ {\cal R}_{\cal C}$, and there is an edge between two vertices if and only if there is a common arc between the two routes in $ {\cal R}_{\cal C}$.
%       
%       Given $u$ and $v$ two vertices of the conflict graph, corresponding to two routes colliding in $ {\cal R}_{\cal C}$. The weight of an edge, $w(u,v)$, is the absolute value of the difference between the distance of the two routes between their respective source node and the collision point.
%       
%       A labeling $F$ of such a graph is an affectation of an integer to each vertex, such that for each vertex $u$, $f(u) \neq f(v)+w(u,v)\mod P$, where $v$ are the neighbors of $u$ in the conflict graph and $P$ our period.

      \includegraphics[scale=0.5]{Fig5.pdf}\\
       \begin{center}
      $\lambda(v,r_i) - \lambda(v,r_j)=1$ 
     \end{center}

      On the other hand if we set all $\lambda(v,r_i) - \lambda(v,r_j)=P$, there is no $P$-periodic affectation if $P<c$.

      \includegraphics[scale=0.5]{Fig6.pdf}\\
       \begin{center}
      $\lambda(v,r_i) - \lambda(v,r_j)=2$

      Here for P=2, there is no $P$-periodic affectation.
      \end{center}

      Therefore if we choose and odd value for $P$ and $c=P+1$, if there is no $P$-periodic affectation but modulo $2$ all $\lambda(v,r_i) - \lambda(v,r_j)$
      are equal to one thus we have a $2$-periodic affectation.

@@

      The main theoretical problem we have to deal with in this context is the following.\\

      \noindent {\bf Problem Periodic Routes Assignment (PRA)}

      \noindent {\bf Input:} a graph $G=(V,A)$, a set $\cal C$ of pair of vertices, a routing function $\cal R$ and an integer $P$.

      \noindent {\bf Question:} does there exist a $P$-periodic affectation of $\cal C$ in $(G,{\cal R})$?

      We will prove in Sec.~\ref{sec:complexity} that the Problem PRA is $\NP$-complete, even in restricted settings.
      In fact, even approximating the smallest value of $P$ for which there is a $P$-periodic assignment is hard.\\
% 
%       \begin{figure}[H]
%       \label{could-ran}
%       \begin{center}
%       % \begin{tabular}{cc}
%       \includegraphics[scale=0.5]{Total-latence.pdf}
%       \caption{Complete process for a leaf in $L$.}
%       \end{center}
%       \end{figure}
%       %\end{tabular}\newline

>>>>

	We consider a directed graph $G=(V,A)$ modelling a network. Each arc  $(u,v)$ in $A$ is labeled by an integer $Dl(u,v) \geq 1$ that we call the delay and
	which represents the number of time slots taken by a signal to go from $u$ to $v$ using this arc. 
	%Note that for any arc $(u,v)$, $Dl(u,v)=Dl(v,u)$. Dominique voulait être plus général, on mettera cette propriété dans notre topologie

@@

      A {\bf routing function} $\cal R$ in $G$ associates to each pair of vertices $(u,v)$ a route from $u$ to $v$. Let $\cal C$ be an {\bf assignment} in $G$, i.e., a set of couples of different vertices of $G$. We denote by $\cal R_{\cal C}$ the set of routes ${\cal R}(u,v)$ for any $(u,v)$ in $\cal C$. We call $\cal R_{\cal C}$ a {\bf routage graph}, it contains all the informations needed in the forthcoming problems (assignment, routes and delays of the arcs).

     \todo{Dire après  la définition des problèmes qu'on pourrait demander de trouver l'assignement et même le routage pour optimiser, mais pas dans cet article
     et qu'on travail avec des réseaux déjà constitués}
      \todo{Si on s'en sert, ajouter ici que le routage est cohérent.}

   \subsection{Slotted time Model}
      Consider now a positive integer $P$ called the {\bf period}. In our problem, we send in the network
      periodic messages of period $P$. The time will thus be cut into slices of $P$ discrete slots. Assume we send a message at the source of the route $r$, at the time slot $m$ in the first period, then a message will be sent at time slot $m$ at each new period. We define the first time slot at which the message reaches a vertex $v$ in this route by $t(v,r) = m + \lambda(v,r) \mod P$.

      A message usually cannot be transported in a single time slot. We denote by $\tau$ the number 
      of consecutive slots necessary to transmit a message. Let us call $[t(v,r)]_{P,\tau}$ the set of time slots used by $r$ at a vertex $v$ in a period $P$, that is $[t(v,r)]_{P,\tau} = \{t(v,r) + i \mod P \mid 0 \leq i < \tau \}$. Usually $P$ and $\tau$ will be clear from the context and we will denote $[t(v,r)]_{P,\tau}$ by $[t(v,r)]$

      A {\bf $(P,\tau)$-periodic affectation} of a routage graph $\cal R$ consists in a set  ${\cal M}=(m_0, \ldots ,m_{c-1})$ of $c$ integers that we call {\bf offsets}, with $c$ the number of routes in $\cal R$. The number $m_i$ represents the index of the first slot used in a period  by the route $r_i \in {\cal R}$ at its source.
      A $P$-periodic affectation must have no {\bf collision} between two routes in ${\cal R}$, that is $\forall (r_i, r_j) \in {\cal R}^2, i \neq j$, % with $\tau$ the size (in number of consecutive slots) of each message that must be periodically sent on each route of ${\cal R}_{\cal C}$, 
      we have $$[t(u,r_i)] \cap [t(u,r_j)] = \emptyset .$$

%       Notice that the notion of $P$-periodic affectation \textbf{is not monotone} with regard to $P$. 
      As an exemple of a $(2,1)$-periodic affectation, let consider a routage graph with routes $\{r_i\}_{i=1,\dots,c}$, such that all pairs of routes intersect at a different edge.
      We set $\tau = 1$ and the delays are chosen so that if $r_i$ and $r_j$ have $v$ as first common vertex then we have $\lambda(v,r_i) - \lambda(v,r_j)=1$.There is a $(2,1)$-periodic affectation by setting all $m_i$ to $0$.

%       A {\bf conflict graph} represents the collision between the routes of $ {\cal R}_{\cal C}$. The vertices of a conflict graph $G = (V,E)$ are the routes of $ {\cal R}_{\cal C}$, and there is an edge between two vertices if and only if there is a common arc between the two routes in $ {\cal R}_{\cal C}$.
%       
%       Given $u$ and $v$ two vertices of the conflict graph, corresponding to two routes colliding in $ {\cal R}_{\cal C}$. The weight of an edge, $w(u,v)$, is the absolute value of the difference between the distance of the two routes between their respective source node and the collision point.
%       
%       A labeling $F$ of such a graph is an affectation of an integer to each vertex, such that for each vertex $u$, $f(u) \neq f(v)+w(u,v)\mod P$, where $v$ are the neighbors of $u$ in the conflict graph and $P$ our period.
      \begin{figure}[h]
       \begin{center}
      \includegraphics[scale=0.7]{Fig5.pdf}
      \end{center}
       \caption{A routage graph with $(0,\dots,0)$ as a $(2,1)$-periodic affectation}
      \end{figure}

@@

    We want to ensure that there is an affectation which allows to send periodic messages from elements in $S$ to elements in $L$.
    The problem we need to solve is thus the following:

      \noindent {\bf  Periodic Routes Assignment (PRA)}

      \noindent {\bf Input:} a routage graph $\cal R$, an integer $\tau$ and an integer $P$.

      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic affectation of $\cal R$ ?

      We will prove in Sec.~\ref{sec:complexity} that the problem PRA is $\NP$-complete, even in restricted settings.
      Even approximating the smallest value of $P$ for which there is a $P$-periodic assignment is hard.
      Another strange property is that given a routage graph, we may have a $(P,\tau)$-periodic affectation but no
      $(P',\tau)$-periodic affectation with $P' > P$, the property is thus not monotone with regards to $P$.

	\begin{lemma} 
	 For any $P$, there is a routage graphe such that there is $(2,1)$-periodic affectation but no $(P,1)$-periodic affectation.
	\end{lemma}
\begin{proof}
      Consider the routage graph ${\cal R}$ given in the previous subsection. 
      We change the delays so that for $v$ the first vertex which belong to $r_i$ and $r_j$,
      we have $\lambda(v,r_i) - \lambda(v,r_j)= P$, where $P$ is an odd number smaller than $c$ the number of routes in ${\cal R}$.
      If we consider a period of $2$, for all $i \neq j$, $\lambda(v,r_i) - \lambda(v,r_j) = 1 $ modulo $2$. Therefore $(0,\dots,0)$ is a $(2,1)$-periodic affectation of ${\cal R}$.
      On the other hand, if we want to find a $(P,1)$-periodic affectation, then $\lambda(v,r_i) - \lambda(v,r_j) = 0 $ modulo $P$.
      Thus we must find $m_i \in \{0,\dots,P-1\}$ such that for all $i \neq j \in \{1,\dots,c\}^2$, $m_i \neq m_j$, which is impossible since $P < c$.

\end{proof}

% 
%       \begin{figure}[H]
%       \label{could-ran}
%       \begin{center}
%       % \begin{tabular}{cc}
%       \includegraphics[scale=0.5]{Total-latence.pdf}
%       \caption{Complete process for a leaf in $L$.}
%       \end{center}
%       \end{figure}
%       %\end{tabular}\newline
